---
title: "Effectiveness of interventions to reduce risky behavior in hominid cryptids"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Write and conduct naive search

Naive search: TS=(("picoides borealis" OR "red-cockaded woodpecker*" OR "red cockaded woodpecker" OR "leuconotopicus borealis" OR woodpecker) AND ("prescribed burn*" OR "prescribed fire*" OR fire* OR wildfire* OR burn*) AND (abundan* OR presen* OR occup* OR occur* OR (popul* NEAR/2 (densit* OR size))))

Searches were conducted on April 10, 2019 with no date restrictions. We searched Scopus (1970-2019) and five databases in Web of Science: the Web of Science Core Collection (1900-2019), BIOSIS Previews (1926-2019), Current Contents Connect (1998-2019), MEDLINE (1950-2019), and Zoological Record (1945-2019).

```{r}
search_directory <- "~/litsearchr/inst/extdata/borealis/"

naiveimport <- litsearchr::import_results(directory = search_directory, remove_duplicates = FALSE, clean_dataset = TRUE, save_full_dataset = FALSE)

dedupe_stage1 <- litsearchr::deduplicate(naiveimport, use_abstracts = FALSE, use_titles = TRUE, method = "quick")
dedupe_stage2 <- litsearchr::deduplicate(dedupe_stage1, use_abstracts = TRUE, use_titles = FALSE, doc_sim = .8, method = "tokens")

naiveresults <- dedupe_stage2

```

## Identify potential keywords

```{r}
rakedkeywords <- litsearchr::extract_terms(naiveresults, type="RAKE", min_freq = 2, title = TRUE, abstract = TRUE, ngrams = TRUE, n=2)
taggedkeywords <- litsearchr::extract_terms(naiveresults, type="tagged", min_freq = 2, title = TRUE, abstract = TRUE, ngrams = TRUE, n=2)

```

## Build the keyword co-occurrence network

```{r}
naivedictionary <- litsearchr::make_dictionary(terms=list(rakedkeywords, taggedkeywords))
naivecorpus <- litsearchr::make_corpus(naiveresults)
naivedfm <- litsearchr::create_dfm(naivecorpus, my_dic=naivedictionary)
naivegraph <- litsearchr::create_network(naivedfm, min_studies=1, min_occurrences = 1)
```

## Identify change points in keyword importance 

```{r}
plot(sort(igraph::strength(BBWO_graph)), ylab="Node strength", main="Ranked node strengths", xlab="Rank")
```

```{r}
splinecutoff <- litsearchr::find_cutoff(naivegraph, method = "spline", degrees = 2, knot_num = 3, 
                                        diagnostics = TRUE, importance_method = "strength")
```


```{r}
reducedgraph <- litsearchr::reduce_graph(naivegraph, cutoff_strength = splinecutoff[1])
searchterms <- litsearchr::get_keywords(reducedgraph, savekeywords = FALSE, makewordle = FALSE)
```

## Group terms into concepts
point counts
woodpecker
```{r}
searchterms
groupedterms <- c()
for(i in 1:length(searchterms)){
print(searchterms[i])
  decision <- menu(choices = c("woodpecker", "fire", "abundance", "none", "multiple"), title = "Which concept group does this term belong to?")
    if(decision==1){
  groupedterms[i] <- "woodpecker"}
  if(decision==2){
  groupedterms[i] <- "fire"
}else if(decision==3){
  groupedterms[i] <- "abundance"
}else if(decision==4){
  groupedterms[i] <- "x"
}else if(decision==5){
  groupedterms[i] <- "multiple"
}
}

term_groups <- cbind(searchterms, groupedterms)

woodpeckers <- unique(append(
  c("woodpecker", "red-cockaded woodpecker", "leuconotopicus borealis", "picoides borealis"),
  term_groups[which(term_groups[,2]=="woodpecker"), 1]))

fire <- unique(append(
  c("fire", "prescribed fire", "prescribed burn", "fire treatment"),
  term_groups[which(term_groups[,2]=="fire"), 1]))

abundance <- unique(append(
  c("species abundance", "species presence", "species density", "population density", "population size"),
  term_groups[which(term_groups[,2]=="abundance"), 1]))

mysearchterms <- list(woodpeckers, fire, abundance)

```

## Write Boolean searches

```{r}
woodpecker_search <- litsearchr::write_search(groupdata = mysearchterms, 
                                          languages = "English", stemming = TRUE,
                                          exactphrase = TRUE, writesearch = FALSE, 
                                          verbose = TRUE)

woodpecker_search

```


## Import search results

```{r}
data1 <- read_bibliography("../inst/extdata/borealis/final_search/bibs/WoS_corecollection_final_1.bib")
data2 <- read_bibliography("../inst/extdata/borealis/final_search/bibs/WoS_corecollection_final_2.bib")
data3 <- read_bibliography("../inst/extdata/borealis/final_search/bibs/scopus_final.bib")

reducing.fields <- c("label","title","author","journal","issn",
                         "volume","pages","year","doi","abstract")

data_all <- rbind(data1[,reducing.fields], data2[,reducing.fields], data3[,reducing.fields])

```


## Deduplicate across databases

```{r}

matches <- find_duplicates(
  data = data_all,
  match_variable = "title",
  group_variable = NULL,
  match_function = "fuzzdist",
  method = "fuzz_m_ratio",
  threshold = 0.1
)

data_unique <- extract_unique_references(data_all, matches)

```


## Screening titles and/or abstracts

```{r}

screen_duplicates(data_unique)

```
