---
title: "Effectiveness of interventions to reduce risky behavior in hominid cryptids"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Write and conduct naive search

Naive search: TS=(("picoides borealis" OR "red-cockaded woodpecker*" OR "red cockaded woodpecker" OR "leuconotopicus borealis" OR woodpecker) AND ("prescribed burn*" OR "prescribed fire*" OR fire* OR wildfire* OR burn*) AND (abundan* OR presen* OR occup* OR occur* OR (popul* NEAR/2 (densit* OR size))))

Searches were conducted on April 10, 2019 with no date restrictions. We searched Scopus (1970-2019) and seven databases in Web of Science: the Web of Science Core Collection (1900-2019), BIOSIS Previews (1926-2019), Current Contents Connect (1998-2019), MEDLINE (1950-2019), Russian Science Citation Index (2005-2019), SciELO Citation Index (1997-2019), and Zoological Record (1945-2019).


```{r}
search_directory <- "~/inst/extdata/borealis/"

naiveimport <- litsearchr::import_results(directory = search_directory, remove_duplicates = FALSE, clean_dataset = TRUE, save_full_dataset = FALSE)

dedupe_stage1 <- litsearchr::deduplicate(naiveimport, use_abstracts = FALSE, use_titles = TRUE, method = "quick")
dedupe_stage2 <- litsearchr::deduplicate(dedupe_stage1, use_abstracts = TRUE, use_titles = FALSE, doc_sim = .8, method = "tokens")

naiveresults <- dedupe_stage2

```

## Identify potential keywords

```{r}
rakedkeywords <- litsearchr::extract_terms(naiveresults, type="RAKE", new_stopwords = NULL, min_freq = 2, title = TRUE, abstract = TRUE, ngrams = TRUE, n=2)
taggedkeywords <- litsearchr::extract_terms(naiveresults, type="tagged", new_stopwords = NULL, min_freq = 2, title = TRUE, abstract = TRUE, ngrams = TRUE, n=2)

```

## Build the keyword co-occurrence network

```{r}
naivedictionary <- litsearchr::make_dictionary(terms=list(rakedkeywords, taggedkeywords))
naivecorpus <- litsearchr::make_corpus(naiveresults)
naivedfm <- litsearchr::create_dfm(naivecorpus, my_dic=naivedictionary)
naivegraph <- litsearchr::create_network(naivedfm, min_studies=2, min_occurrences = 2)
```

## Identify change points in keyword importance 

```{r}
plot(sort(igraph::strength(BBWO_graph)), ylab="Node strength", main="Ranked node strengths", xlab="Rank")
```

```{r}
splinecutoff <- litsearchr::find_cutoff(naivegraph, method = "spline", degrees = 2, knot_num = 3, 
                                        diagnostics = TRUE, importance_method = "strength")
```


```{r}
reducedgraph <- litsearchr::reduce_graph(naivegraph, cutoff_strength = splinecutoff[1])
searchterms <- litsearchr::get_keywords(reducedgraph, savekeywords = FALSE, makewordle = FALSE)
```

## Group terms into concepts

```{r}

```

## Write Boolean searches

```{r}
litsearchr::usable_databases()
```

```{r}
bigfootsearch <- litsearchr::write_search(groupdata = mysearchterms, 
                                          languages = "English", stemming = TRUE,
                                          exactphrase = TRUE, writesearch = FALSE, 
                                          verbose = TRUE)

bigfootsearch
```
